Scheeper's Meager-NWD Game and the Menger Game

### 2014-08-25 phd/dissertation/chapter3.tex
Name and camelCase based game commands/definitions

\begin{game}
  Let $\congame{X}{S}$ denote the \term{$W$-convergence game} with
  players $\pl O$, $\pl P$, for a topological space $X$ and $S\subseteq X$.

@@

  If $S=\{x\}$ then we write $\congame{X}{x}$ for short.
\end{game}

@@

\begin{defn}
  The spaces $X$ for which $\pl O \win \congame{X}{x}$ for all $x\in X$ are
  called \term{$W$-spaces}.
\end{defn}

@@

\begin{prop}
  $X$ is first countable
    if and only if
  $\pl O \prewin \congame{X}{x}$ for all $x\in X$.
\end{prop}

@@

  If $X$ is not first countable at some $x$, let $\sigma$ be a
  predetermined strategy for $\pl O$ in $\congame{X}{x}$. There exists
  an open neighborhood $U$ of $x$ which does not contain any
  $\bigcap_{m\leq n}\sigma(m)$ (otherwise
  $\{\bigcap_{m\leq n}\sigma(m):n<\omega\}$ would be a countable base at $x$).
  Let $x_n$ be an element of $\bigcap_{m\leq n}\sigma(m)\setminus U$ for all
  $n<\omega$. Then $\<x_0,x_1,\dots\>$ is a winning counter-attack to $\sigma$
  for $\pl P$, so $\pl O$ lacks a winning predetermined strategy.
\end{proof}

At first glance, the difficulty of $\congame{X}{S}$ could be increased for
$\pl O$ by only restricting the choices for $\pl P$ to be within the most
recent open set played by $\pl O$, rather than all the previously played
open sets.

\begin{defn}
  Let $\congamehard{X}{S}$ denote the \term{hard $W$-convergence game} which
  proceeds as $\congame{X}{S}$, except that $\pl P$ need only choose
  $x_n\in O_n$ rather than $x_n\in \bigcap_{m\leq n}O_m$ during each round.
\end{defn}

@@

\begin{prop}\label{propNormalHardCon}
  $\pl O \limitwin \congame{X}{S}$
    if and only if
  $\pl O \limitwin \congamehard{X}{S}$,
  where $\limitwin$ is either $\win$ or $\prewin$.
\end{prop}

@@

\begin{game}
  Let $\clusgame{X}{S}$ ($\clusgamehard{X}{S}$) be a variation of
  $\congame{X}{S}$ ($\congamehard{X}{S}$) such that $x_n$ need only
  cluster at $S$, that is, for every open neighborhood $U$ of $S$, $x_n\in U$
  for infinitely many $n<\omega$.
\end{game}

@@

\begin{prop}
  $\pl O \limitwin \congame{X}{S}$
    if and only if
  $\pl O \limitwin \clusgame{X}{S}$
  where $\limitwin$ is any of $\win$, $\prewin$, $\tactwin$, or $\markwin$.
\end{prop}

@@

  In the predetermined (resp. tactical) case, suppose that $\sigma$ is a
  winning predetermined (resp. tactical) strategy for $\pl O$ in
  $\clusgame{X}{S}$. Let $p$ be a legal attack against $\sigma$, and $q$ be a
  subsequence of $p$. It's easily seen that $q$ is also a legal attack against
  $\sigma$, so $q$ clusters at $S$. Since every subsequence of $p$ clusters
  at $S$, $p$ converges to $S$, and $\sigma$ is a winning predetermined
  (resp. tactical) strategy for $\pl O$ in $\congame{X}{S}$ as well.

  In the final case, note that any Markov strategy $\sigma'$ for $\pl O$ may
  be strengthened to $\sigma$ defined by
  $\sigma(x,n)=\bigcap_{m\leq n}\sigma'(x,m)$.
  So, suppose that $\sigma$ is a winning Markov strategy for $\pl O$ in
  $\clusgame{X}{S}$ such that $\sigma(x,m)\supseteq\sigma(x,n)$ for all
  $m\leq n$.

  Let $p$ be a legal attack against $\sigma$, and $q$ be a subsequence of $p$.
  For $m<\omega$, there exists $f(m)\geq m$ such that $q(m)=p(f(m))$. It follows
  that
    $
      q(0)=p(f(0))
        \in
      \sigma(\emptyset,0)\cap \bigcap_{m\leq f(0)}\sigma(\<p(m)\>,m)
        \subseteq
      \sigma(\emptyset,0)
    $
  and
    \[
      \begin{array}{rcl}
        q(n+1)=p(f(n+1)) &
          \in &
        \displaystyle
        \sigma(\emptyset,0)\cap \bigcap_{m<f(n+1)}\sigma(\<p(m)\>,m+1) \\ &
          \subseteq &
        \displaystyle
        \sigma(\emptyset,0)\cap \bigcap_{m<n+1}\sigma(\<p(f(m))\>,f(m)+1) \\ &
          = &
        \displaystyle
        \sigma(\emptyset,0)\cap \bigcap_{m<n+1}\sigma(\<q(m)\>,f(m)+1) \\ &
          \subseteq &
        \displaystyle
        \sigma(\emptyset,0)\cap \bigcap_{m<n+1}\sigma(\<q(m)\>,m+1)
      \end{array}
    \]
  so $q$ is also a legal attack against $\sigma$. Since $\sigma$ is a winning
  strategy, $q$ clusters at $S$, and since every subsequence of $p$ clusters
  at $S$, $p$ must converge to $S$. Thus $\sigma$ is also a winning Markov
  strategy for $\pl O$ in $\congame{X}{S}$ as well.
\end{proof}

@@

\begin{ques}
  Does $\pl O \ktactwin{2} \clusgame{X}{S}$ imply
  $\pl O \ktactwin{2} \congame{X}{S}$? What about for $\kmarkwin{2}$?
\end{ques}

\begin{ques}
  Could $\pl O \ktactwin{k+1} \congame{X}{S}$ actually imply
  $\pl O \tactwin \congame{X}{S}$? What about for $\clusgame{X}{S}$?
\end{ques}

@@

\begin{prop}\label{propFortTact}
  $\pl O \tactwin \congame{\oneptcomp\kappa}{\infty}$ for all cardinals $\kappa$
\end{prop}

@@

\begin{cor}
  $\pl O \win \congamehard{\oneptcomp\kappa}{\infty}$ for all cardinals $\kappa$
\end{cor}

@@

Since it's trivial to show that $\pl O\prewin\congame{\oneptcomp\kappa}$ if and
only if $\kappa\leq\omega$, this closes the question on limited information
strategies for $\congame{\oneptcomp\kappa}{\infty}$. However, limited
information analysis of the harder $\congamehard{\oneptcomp\kappa}{\infty}$
is more interesting.

@@

\begin{thm}
  $\pl O \not\markwin \congamehard{\oneptcomp\omega_1}{\infty}$.
\end{thm}

@@

\begin{thm}
  $\pl O \not\kmarkwin{k} \congamehard{\oneptcomp\omega_1}{\infty}$.
\end{thm}

@@

\begin{thm}
  $\pl O\markwin \clusgamehard{\oneptcomp\omega_1}{\infty}$.
\end{thm}

@@

\begin{thm}
  $O\not\ktactwin{k}\clusgamehard{\oneptcomp\omega_1}{\infty}$.
\end{thm}

\begin{proof}
  Let $\sigma$ be a $k$-tactic for $\pl O$ in
  $\clusgamehard{\oneptcomp\omega_1}{\infty}$. By the closing-up lemma, the set
    \[
      C_\sigma
        =
      \{
        \alpha<\omega_1
          :
        s\in\alpha^{\leq k}
          \Rightarrow
        \oneptcomp\omega_1\setminus\sigma(s)
        \subset \alpha
      \}
    \]
  is closed and unbounded. Let $a_\sigma:\omega_1\to C_\sigma$ be an order
  isomorphism.

@@

\begin{thm}
  $O\not\kmarkwin{k}\clusgamehard{\oneptcomp\omega_2}{\infty}$.
\end{thm}

\begin{proof}
  Let $\sigma$ be a $k$-mark for $\pl O$ in
  $\clusgame{\oneptcomp\omega_2}{\infty}$. By the closing-up lemma, the set
    \[
      C_\sigma
        =
      \{
        \alpha<\omega_2
          :
        s\in\alpha^{<\omega}
          \Rightarrow
        \oneptcomp\omega_2\setminus\sigma\circ\mu_k(s)
        \subset \alpha
      \}
    \]
  is closed and unbounded. (Recall that $\mu_k$ is the
  $k$-Markov fog-of-war which turns perfect information into the last $k$
  moves and the round number.) Let $a_\sigma:\omega_2\to C_\sigma$ be
  an order isomorphism.

@@

\begin{prop}
  Suppose $S\subseteq Y\subseteq X$, $\limitwin$ is any of $\win$,
  $\ktactwin{k}$, or $\kmarkwin{k}$, and $G(X,S)$ is any of $\congame{X}{S}$,
  $\congamehard{X}{S}$, $\clusgame{X}{S}$, or $\clusgamehard{X}{S}$.

@@

\begin{cor}
  $\pl O\not\ktactwin{k}\clusgamehard{\SigmaProdR{\omega_1}}{\vec0}$,
  $\pl O\not\kmarkwin{k}\congamehard{\SigmaProdR{\omega_1}}{\vec0}$, and
  $\pl O\not\kmarkwin{k}\clusgamehard{\SigmaProdR{\omega_2}}{\vec0}$.
\end{cor}

@@

\begin{prop}
  $\pl O \codewin \congame{\oneptcomp\kappa}{\infty}$.
\end{prop}

@@

\begin{game}
  Let $\pfgame\kappa$ denote the \term{point-finite game} with players
  $\pl F$, $\pl C$ for each cardinal $\kappa$.

@@

\begin{lem}
  $\pl F \codewin \pfgame\kappa$ implies
  $\pl O \codewin \congame{\SigmaProdR\kappa}{\vec0}$.
\end{lem}

\begin{proof}
  Let $\sigma$ be a winning coding strategy for $\pl F$ in $\pfgame\kappa$
  such that $\sigma(\emptyset)\supset\emptyset$ and $\sigma(F,C)\supset F$.

@@

  We define the coding strategy $\tau$ for $\pl O$ in
  $\congame{\SigmaProdR\kappa}{\vec0}$ as follows:
  \[
    \tau(\emptyset)
      =
    U(\sigma(\emptyset),1) \cap \SigmaProdR\kappa
  \]
  \[
    \tau(\<U(F,\epsilon)\cap\SigmaProdR\kappa,x\>)
      =
    U\left(\sigma(\<F,C_{\epsilon}(x)\>),\frac{\epsilon}{2}\right)
      \cap
    \SigmaProdR\kappa
  \]

@@

\begin{thm}
  $\pl F \codewin \pfgame{\kappa}$ for all cardinals $\kappa$.
\end{thm}

\begin{proof}
  For each cardinal $\kappa$ and $\lambda<\kappa$,
  assume $\sigma_\lambda$ is a winning strategy for $\pl F$ in $\pfgame\lambda$
  such that $\sigma_\lambda(\emptyset)\supset\emptyset$ and
  $\sigma_\lambda(\<F,C\>)\supset F$.

@@

\begin{cor}
  $\pl O\codewin\congame{\SigmaProdR{\kappa}}{\vec{0}}$
  for all cardinals $\kappa$.
\end{cor}

@@

\begin{ques}
  Does $\pl O \win\congame{X}{x}$ imply $\pl O \codewin\congame{X}{x}$?
\end{ques}

>>>>

\begin{game}
  Let $\gruConGame{X}{S}$ denote the \term{$W$-convergence game} with
  players $\pl O$, $\pl P$, for a topological space $X$ and $S\subseteq X$.

@@

  If $S=\{x\}$ then we write $\gruConGame{X}{x}$ for short.
\end{game}

@@

\begin{defn}
  The spaces $X$ for which $\pl O \win \gruConGame{X}{x}$ for all $x\in X$ are
  called \term{$W$-spaces}.
\end{defn}

@@

\begin{prop}
  $X$ is first countable
    if and only if
  $\pl O \prewin \gruConGame{X}{x}$ for all $x\in X$.
\end{prop}

@@

  If $X$ is not first countable at some $x$, let $\sigma$ be a
  predetermined strategy for $\pl O$ in $\gruConGame{X}{x}$. There exists
  an open neighborhood $U$ of $x$ which does not contain any
  $\bigcap_{m\leq n}\sigma(m)$ (otherwise
  $\{\bigcap_{m\leq n}\sigma(m):n<\omega\}$ would be a countable base at $x$).
  Let $x_n$ be an element of $\bigcap_{m\leq n}\sigma(m)\setminus U$ for all
  $n<\omega$. Then $\<x_0,x_1,\dots\>$ is a winning counter-attack to $\sigma$
  for $\pl P$, so $\pl O$ lacks a winning predetermined strategy.
\end{proof}

At first glance, the difficulty of $\gruConGame{X}{S}$ could be increased for
$\pl O$ by only restricting the choices for $\pl P$ to be within the most
recent open set played by $\pl O$, rather than all the previously played
open sets.

\begin{defn}
  Let $\gruConGameHard{X}{S}$ denote the \term{hard $W$-convergence game} which
  proceeds as $\gruConGame{X}{S}$, except that $\pl P$ need only choose
  $x_n\in O_n$ rather than $x_n\in \bigcap_{m\leq n}O_m$ during each round.
\end{defn}

@@

\begin{prop}\label{propNormalHardCon}
  $\pl O \limitwin \gruConGame{X}{S}$
    if and only if
  $\pl O \limitwin \gruConGameHard{X}{S}$,
  where $\limitwin$ is either $\win$ or $\prewin$.
\end{prop}

@@

\begin{game}
  Let $\gruClusGame{X}{S}$ ($\gruClusGameHard{X}{S}$) be a variation of
  $\gruConGame{X}{S}$ ($\gruConGameHard{X}{S}$) such that $x_n$ need only
  cluster at $S$, that is, for every open neighborhood $U$ of $S$, $x_n\in U$
  for infinitely many $n<\omega$.
\end{game}

@@

\begin{prop}
  $\pl O \limitwin \gruConGame{X}{S}$
    if and only if
  $\pl O \limitwin \gruClusGame{X}{S}$
  where $\limitwin$ is any of $\win$, $\prewin$, $\tactwin$, or $\markwin$.
\end{prop}

@@

  In the predetermined (resp. tactical) case, suppose that $\sigma$ is a
  winning predetermined (resp. tactical) strategy for $\pl O$ in
  $\gruClusGame{X}{S}$. Let $p$ be a legal attack against $\sigma$, and $q$ be a
  subsequence of $p$. It's easily seen that $q$ is also a legal attack against
  $\sigma$, so $q$ clusters at $S$. Since every subsequence of $p$ clusters
  at $S$, $p$ converges to $S$, and $\sigma$ is a winning predetermined
  (resp. tactical) strategy for $\pl O$ in $\gruConGame{X}{S}$ as well.

  In the final case, note that any Markov strategy $\sigma'$ for $\pl O$ may
  be strengthened to $\sigma$ defined by
  $\sigma(x,n)=\bigcap_{m\leq n}\sigma'(x,m)$.
  So, suppose that $\sigma$ is a winning Markov strategy for $\pl O$ in
  $\gruClusGame{X}{S}$ such that $\sigma(x,m)\supseteq\sigma(x,n)$ for all
  $m\leq n$.

  Let $p$ be a legal attack against $\sigma$, and $q$ be a subsequence of $p$.
  For $m<\omega$, there exists $f(m)\geq m$ such that $q(m)=p(f(m))$. It follows
  that
    $
      q(0)=p(f(0))
        \in
      \sigma(\emptyset,0)\cap \bigcap_{m\leq f(0)}\sigma(\<p(m)\>,m)
        \subseteq
      \sigma(\emptyset,0)
    $
  and
    \[
      \begin{array}{rcl}
        q(n+1)=p(f(n+1)) &
          \in &
        \displaystyle
        \sigma(\emptyset,0)\cap \bigcap_{m<f(n+1)}\sigma(\<p(m)\>,m+1) \\ &
          \subseteq &
        \displaystyle
        \sigma(\emptyset,0)\cap \bigcap_{m<n+1}\sigma(\<p(f(m))\>,f(m)+1) \\ &
          = &
        \displaystyle
        \sigma(\emptyset,0)\cap \bigcap_{m<n+1}\sigma(\<q(m)\>,f(m)+1) \\ &
          \subseteq &
        \displaystyle
        \sigma(\emptyset,0)\cap \bigcap_{m<n+1}\sigma(\<q(m)\>,m+1)
      \end{array}
    \]
  so $q$ is also a legal attack against $\sigma$. Since $\sigma$ is a winning
  strategy, $q$ clusters at $S$, and since every subsequence of $p$ clusters
  at $S$, $p$ must converge to $S$. Thus $\sigma$ is also a winning Markov
  strategy for $\pl O$ in $\gruConGame{X}{S}$ as well.
\end{proof}

@@

\begin{ques}
  Does $\pl O \ktactwin{2} \gruClusGame{X}{S}$ imply
  $\pl O \ktactwin{2} \gruConGame{X}{S}$? What about for $\kmarkwin{2}$?
\end{ques}

\begin{ques}
  Could $\pl O \ktactwin{k+1} \gruConGame{X}{S}$ actually imply
  $\pl O \tactwin \gruConGame{X}{S}$? What about for $\gruClusGame{X}{S}$?
\end{ques}

@@

\begin{prop}\label{propFortTact}
  $\pl O \tactwin \gruConGame{\oneptcomp\kappa}{\infty}$ for all cardinals $\kappa$
\end{prop}

@@

\begin{cor}
  $\pl O \win \gruConGameHard{\oneptcomp\kappa}{\infty}$ for all cardinals $\kappa$
\end{cor}

@@

Since it's trivial to show that $\pl O\prewin\gruConGame{\oneptcomp\kappa}$ if and
only if $\kappa\leq\omega$, this closes the question on limited information
strategies for $\gruConGame{\oneptcomp\kappa}{\infty}$. However, limited
information analysis of the harder $\gruConGameHard{\oneptcomp\kappa}{\infty}$
is more interesting.

@@

\begin{thm}
  $\pl O \not\markwin \gruConGameHard{\oneptcomp\omega_1}{\infty}$.
\end{thm}

@@

\begin{thm}
  $\pl O \not\kmarkwin{k} \gruConGameHard{\oneptcomp\omega_1}{\infty}$.
\end{thm}

@@

\begin{thm}
  $\pl O\markwin \gruClusGameHard{\oneptcomp\omega_1}{\infty}$.
\end{thm}

@@

\begin{thm}
  $O\not\ktactwin{k}\gruClusGameHard{\oneptcomp\omega_1}{\infty}$.
\end{thm}

\begin{proof}
  Let $\sigma$ be a $k$-tactic for $\pl O$ in
  $\gruClusGameHard{\oneptcomp\omega_1}{\infty}$. By the closing-up lemma, the set
    \[
      C_\sigma
        =
      \{
        \alpha<\omega_1
          :
        s\in\alpha^{\leq k}
          \Rightarrow
        \oneptcomp\omega_1\setminus\sigma(s)
        \subset \alpha
      \}
    \]
  is closed and unbounded. Let $a_\sigma:\omega_1\to C_\sigma$ be an order
  isomorphism.

@@

\begin{thm}
  $O\not\kmarkwin{k}\gruClusGameHard{\oneptcomp\omega_2}{\infty}$.
\end{thm}

\begin{proof}
  Let $\sigma$ be a $k$-mark for $\pl O$ in
  $\gruClusGame{\oneptcomp\omega_2}{\infty}$. By the closing-up lemma, the set
    \[
      C_\sigma
        =
      \{
        \alpha<\omega_2
          :
        s\in\alpha^{<\omega}
          \Rightarrow
        \oneptcomp\omega_2\setminus\sigma\circ\mu_k(s)
        \subset \alpha
      \}
    \]
  is closed and unbounded. (Recall that $\mu_k$ is the
  $k$-Markov fog-of-war which turns perfect information into the last $k$
  moves and the round number.) Let $a_\sigma:\omega_2\to C_\sigma$ be
  an order isomorphism.

@@

\begin{prop}
  Suppose $S\subseteq Y\subseteq X$, $\limitwin$ is any of $\win$,
  $\ktactwin{k}$, or $\kmarkwin{k}$, and $G(X,S)$ is any of $\gruConGame{X}{S}$,
  $\gruConGameHard{X}{S}$, $\gruClusGame{X}{S}$, or $\gruClusGameHard{X}{S}$.

@@

\begin{cor}
  $\pl O\not\ktactwin{k}\gruClusGameHard{\SigmaProdR{\omega_1}}{\vec0}$,
  $\pl O\not\kmarkwin{k}\gruConGameHard{\SigmaProdR{\omega_1}}{\vec0}$, and
  $\pl O\not\kmarkwin{k}\gruClusGameHard{\SigmaProdR{\omega_2}}{\vec0}$.
\end{cor}

@@

\begin{prop}
  $\pl O \codewin \gruConGame{\oneptcomp\kappa}{\infty}$.
\end{prop}

@@

\begin{game}
  Let $\cloPFGame\kappa$ denote the \term{point-finite game} with players
  $\pl F$, $\pl C$ for each cardinal $\kappa$.

@@

\begin{lem}
  $\pl F \codewin \cloPFGame\kappa$ implies
  $\pl O \codewin \gruConGame{\SigmaProdR\kappa}{\vec0}$.
\end{lem}

\begin{proof}
  Let $\sigma$ be a winning coding strategy for $\pl F$ in $\cloPFGame\kappa$
  such that $\sigma(\emptyset)\supset\emptyset$ and $\sigma(F,C)\supset F$.

@@

  We define the coding strategy $\tau$ for $\pl O$ in
  $\gruConGame{\SigmaProdR\kappa}{\vec0}$ as follows:
  \[
    \tau(\emptyset)
      =
    U(\sigma(\emptyset),1) \cap \SigmaProdR\kappa
  \]
  \[
    \tau(\<U(F,\epsilon)\cap\SigmaProdR\kappa,x\>)
      =
    U\left(\sigma(\<F,C_{\epsilon}(x)\>),\frac{\epsilon}{2}\right)
      \cap
    \SigmaProdR\kappa
  \]

@@

\begin{thm}
  $\pl F \codewin \cloPFGame{\kappa}$ for all cardinals $\kappa$.
\end{thm}

\begin{proof}
  For each cardinal $\kappa$ and $\lambda<\kappa$,
  assume $\sigma_\lambda$ is a winning strategy for $\pl F$ in $\cloPFGame\lambda$
  such that $\sigma_\lambda(\emptyset)\supset\emptyset$ and
  $\sigma_\lambda(\<F,C\>)\supset F$.

@@

\begin{cor}
  $\pl O\codewin\gruConGame{\SigmaProdR{\kappa}}{\vec{0}}$
  for all cardinals $\kappa$.
\end{cor}

@@

\begin{ques}
  Does $\pl O \win\gruConGame{X}{x}$ imply $\pl O \codewin\gruConGame{X}{x}$?
\end{ques}
